A Type System for String Sanitation\\Implemented Inside a Python

### 2014-04-16 sanitation-plas14/sanitation-plas14.tex
P -> T, but macros still say T :-(

% Judgements
\newcommand{\err}{{\tt err}}
\newcommand{\trden}[1]{[[#1]]} % = Translation Denotation.

@@

<T> ::= \hfill Internal expressions \\
  $\tstr{T}$ \alt
  $\siconcat{T}{T}$ \alt
  $\tsubst{r}{T}{T}$ \alt
  $\tcheck{r}{T}$
\end{grammar}
\caption{Syntax for the string sanitation fragment of $\lcs$}
\label{fig:lcsSyntax}
\end{figure}

@@

\inferline{T-T-String}
{\tctx \vdash \tstr{s} : \str}
{\cdot}

\inferline{T-T-Concat}
{\tctx \vdash \tconcat{T_1}{T_2} : \str}
{\tctx \vdash T_1 : \str \\ \tctx \vdash T_2 : \str}

\inferline{T-T-Subst}
{\tctx \vdash \tsubst{\rx{r}}{T_1}{T_2} : \str}
{\tctx \vdash T_1 : \str \\ \tctx \vdash T_2 : \str}

\inferline{T-T-Check}
{\tctx \vdash \tcheck{\rx{r}}{s} : \str}
{\tctx \vdash T : \str}

\end{tabular}
\caption{Typing rules for target langauge terms (T). The metavariable s ranges over
string literals.}
\label{fig:etyping}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\begin{tabular}{c r}
\inferline{T-E-Str}
{\tstr{s} \treduces \str{s}}
{ \cdot }

\inferline{T-E-Include}
{\rx{r} \treduces \rx{r}}
{ \cdot }

\inferline{T-E-Concat}
{\tconcat{T_1}{T_2} \treduces \tstr{s_1 \cdot s_2}} % ???
{T_1 \treduces \tstr{s_1} \\ T_2 \treduces \tstr{s_2}}

\inferline{T-E-Subst}
{\tsubst{r}{T_2}{T_3} \treduces \tstr{s}}
{T_1 \treduces \rx{r} \\ T_2 \treduces \tstr{s_2} \\ T_3 \treduces \tstr{s_3} \\ \dosubst{r}{s_2}{s_3} = s}

\inferline{T-E-Check-OK}
{\tcheck{r}{T} \treduces \tstr{s}}
{T \treduces \strin{s} \\ s \in \lang{r}}

\inferline{T-E-Check-NotOk}
{T \err}
{\tcheck{r}{T} \treduces \tstr{s} \\ s \not \in \lang{r}}

@@

\inferline{Tr-Concat}
{\trden{\siconcat{S_1}{S_2}} = \tconcat{T_1}{T_2}}
{ \trden{S_1} = T_1 \\ \trden{S_2} = T_2}

\inferline{Tr-Subst}
{\trden{ \sisubst{r}{S_1}{S_2} } = \tsubst{\rx{r}}{T_1}{T_2} }
{ \trden{S_1} = T_1 \\ \trden{S_2} = T_2 }

\inferline{Tr-Subst}
{\trden{ \coerce{r'}{S} } = T}
{ \trden{s} = T \\ S : \strin{r} \\ \lang{r'} \subseteq \lang{r}}

\inferline{Tr-Coerce-NotOk}
{\trden{\coerce{r'}{S}} = \tcheck{\rx{r'}}{S}}
{\trden{S} = T \\ s : \strin{r} \\ \lang{r'} \not \subseteq \lang{r}}

\end{tabular}
\caption{Translation from source terms (S) to target terms (T).
The translation is type-directed in the Tr-Coerce cases.}
\end{center}
\end{figure}

@@

Note that we never insert an internal check where the type of the string implies
that a check must succeed. Furthermore, expensive calculations (such as language
inclusion or ${\tt computer\_replacement\_language}$, all occur at compile time.
Since sanitation problems are generally expressible with smalle, simple expressions,
we believe that the compile-time overhead is not significant enough to prohibiour target usecase. Informal experimentation with our implementation seems to
support this assertion.

>>>>

% Judgements
\newcommand{\err}{{\tt err}}
\newcommand{\trden}[1]{\llbracket #1 \rrbracket} % = Translation Denotation.

@@

<P> ::= \hfill Internal expressions \\
  $\tstr{P}$ \alt
  $\siconcat{P}{P}$ \alt
  $\tsubst{r}{P}{P}$ \alt
  $\tcheck{r}{P}$
\end{grammar}
\caption{Syntax for the string sanitation fragment of $\lcs$}
\label{fig:lcsSyntax}
\end{figure}

@@

\inferline{P-T-String}
{\tctx \vdash \tstr{s} : \str}
{\cdot}

\inferline{P-T-Concat}
{\tctx \vdash \tconcat{P_1}{P_2} : \str}
{\tctx \vdash P_1 : \str \\ \tctx \vdash P_2 : \str}

\inferline{P-T-Subst}
{\tctx \vdash \tsubst{\rx{r}}{P_1}{P_2} : \str}
{\tctx \vdash P_1 : \str \\ \tctx \vdash P_2 : \str}

\inferline{P-T-Check}
{\tctx \vdash \tcheck{\rx{r}}{s} : \str}
{\tctx \vdash P : \str}

\end{tabular}
\caption{Typing rules for target langauge terms (P). The metavariable s ranges over
string literals.}
\label{fig:etyping}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\begin{tabular}{c r}
\inferline{P-E-Str}
{\tstr{s} \treduces \str{s}}
{ \cdot }

\inferline{P-E-Include}
{\rx{r} \treduces \rx{r}}
{ \cdot }

\inferline{P-E-Concat}
{\tconcat{P_1}{P_2} \treduces \tstr{s_1 \cdot s_2}} % ???
{P_1 \treduces \tstr{s_1} \\ P_2 \treduces \tstr{s_2}}

\inferline{P-E-Subst}
{\tsubst{r}{P_2}{P_3} \treduces \tstr{s}}
{P_1 \treduces \rx{r} \\ P_2 \treduces \tstr{s_2} \\ P_3 \treduces \tstr{s_3} \\ \dosubst{r}{s_2}{s_3} = s}

\inferline{P-E-Check-OK}
{\tcheck{r}{P} \treduces \tstr{s}}
{P \treduces \strin{s} \\ s \in \lang{r}}

\inferline{P-E-Check-NotOk}
{P \err}
{\tcheck{r}{T} \treduces \tstr{s} \\ s \not \in \lang{r}}

@@

\inferline{Tr-Concat}
{\trden{\siconcat{S_1}{S_2}} = \tconcat{P_1}{P_2}}
{ \trden{S_1} = P_1 \\ \trden{S_2} = P_2}

\inferline{Tr-Subst}
{\trden{ \sisubst{r}{S_1}{S_2} } = \tsubst{\rx{r}}{P_1}{P_2} }
{ \trden{S_1} = P_1 \\ \trden{S_2} = P_2 }

\inferline{Tr-Subst}
{\trden{ \coerce{r'}{S} } = P}
{ \trden{s} = P \\ S : \strin{r} \\ \lang{r'} \subseteq \lang{r}}

\inferline{Tr-Coerce-NotOk}
{\trden{\coerce{r'}{S}} = \tcheck{\rx{r'}}{S}}
{\trden{S} = P \\ s : \strin{r} \\ \lang{r'} \not \subseteq \lang{r}}

\end{tabular}
\caption{Translation from source terms (S) to target terms (P).
The translation is type-directed in the Tr-Coerce cases.}
\end{center}
\end{figure}

@@

Note that we never insert an internal check where the type of the string implies
that a check must succeed. Furthermore, expensive calculations (such as language
inclusion or ${\tt dsubst}$, all occur at compile time.
Since sanitation problems are generally expressible with smalle, simple expressions,
we believe that the compile-time overhead is not significant enough to prohibiour target usecase. Informal experimentation with our implementation seems to
support this assertion.
